---
output:
  word_document: default
---
---
title: "Perinatal"
author: "Yuster"
date: "1/7/2018"
output: word_document
--
```{r setup, include=FALSE}
library(dplyr)
library(tidyr)
library(tidyverse)
library(tidyselect)
library(readr)
```

## r notes
The report of perinatal death in Nakuru County from 2014 to 2017 dataset.
Perinata Death review forms are filled in the facility where the event occured and uploaded to DHIS2.Of more than 2000 events are available from the system. The dataset were downloaded and data cleaning done to remove data elements that does not fit definition of perinatal death, multiple entries and duplicates were deleted as well as empty columns.
```{r dataset, include=FALSE}
perdatn <- readr::read_csv("~/Dropbox/Peri2018.csv")
```

```{r glimpse, include=FALSE}
glimpse(perdatn)
```

```{r, prop}
perdatn %>% 
  with(table(Org_unit, Year)) %>%
  prop.table(margin=2)*100
```

```{r, org-unit}
perdatn %>% ## reported no of perinatal deaths by year in 2015 we had 431  deaths
  group_by(Org_unit) %>% 
  summarise(count = n()) %>% 
  mutate(rfreq = count/sum(count)*100) %>% 
  mutate_if(is.numeric, funs(round(., 0))) %>% ## relative frequency 
  arrange(desc(rfreq))
```

## rmarkdown
Babies delivered by Nulliparous women are at greater risk of perinatal risk than multiparous women.The dataset is recorded on number of gravida mother ever had from 22 weeks gestation including the current pregnancy.

```{r recodegravida, include=FALSE}
perdatn$Mgravrec <- case_when(perdatn$M_gravida <= 1 ~ 'Nulliparous',
                         perdatn$M_gravida >=2 ~ 'Multiparous')
```
```{r, gravida}
perdatn %>% 
  group_by(Mgravrec) %>% 
  summarise(count=n()) %>% 
  mutate(rfreq= count/sum(count)*100) %>%
  mutate_if(is.numeric, funs(round(., 1))) %>% 
  arrange(desc(rfreq))
```
```{r, recode gestation, include=FALSE}
perdatn$rmgest <- case_when(perdatn$M_gest <= 28 ~ 'Ext preterm',
                       between(perdatn$M_gest, 28, 32) ~ 'Very preterm',
                       between(perdatn$M_gest, 33, 37) ~ 'Late preterm',
                       between(perdatn$M_gest, 38, 40) ~ ' Full term',
                       perdatn$M_gest >= 41 ~ 'post term')
```
## Gestational Weeks
Using WHO classification of preterm, gestational weeks were classified on extremely preterm (Less than 28 weeks), very preterm(28-32 weeks)
moderate to late preterm(32-37weeks) and post mature from 41 weeks and above. From the dataset 794 varibales were complete with mean gestational
of 34.5+/- 5.43, Min of 18 weeks and max of 48 weeks with skew of -0.59 and kurtosis of -0.45.

```{r, rgestation}
perdatn %>% 
  group_by(rmgest) %>% 
  summarise(count=n()) %>% 
  mutate(rfreq= count/sum(count)*100) %>% 
  mutate_if(is.numeric, funs(round(., 1))) %>% 
  arrange(desc(rfreq))
```
## Type of pregnancy
A total of 765 (82%) of the cases were singleton, less than 15% being twin or triplet.

```{r, pregtype}
perdatn %>% 
  group_by(preg_type) %>% 
  summarise(count=n()) %>% 
  mutate(rfreq= count/sum(count)*100) %>% 
  mutate_if(is.numeric, funs(round(., 1))) %>% 
  arrange(desc(rfreq))
```
## Presentation of foetus
Cephalic account for 77% of the perinatal death

```{r, fpresen}
perdatn %>% 
  group_by(foetus_pres) %>% 
  summarise(count=n()) %>% 
  mutate(rfreq= count/sum(count)*100) %>% 
  mutate_if(is.numeric, funs(round(., 1))) %>% 
  arrange(desc(rfreq))
```
## Time of Newborn death
Of the deaths, 60% (557/929) died with 7 days of live.
```{r, timeof newborn}
perdatn %>% 
  group_by(t_nb_death) %>% 
  summarise(count=n()) %>% 
  mutate(rfreq= count/sum(count)*100) %>% 
  mutate_if(is.numeric, funs(round(., 1))) %>% 
  arrange(desc(rfreq))
```







